\section{Methods}
\label{sec:methods}

%==========================================================================
% TODO: PÁRRAFO INICIAL - Resumen de la sección
%==========================================================================
% [AQUÍ FALTA: Un párrafo breve (2-3 oraciones) que resuma la estrategia
% metodológica y anticipe las subsecciones. En NeurIPS esto ayuda al revisor.
%
``To systematically evaluate the circuit mechanisms underlying extra-classical
surround suppression, we implemented a large-scale spiking neural network
using the NEST simulator \citep{Gewaltig2007}. We constructed three
architectural configurations of increasing complexity, all grounded in the
canonical cortical microcircuit \citep{Potjans2014}: (1) V1 columns with only
lateral intra-areal connections, (2) a V1-V2 hierarchy with simple convergent
feedback, and (3) a V1-V2 hierarchy with reciprocal lateral inhibition between
V2 modules. Below we describe the neuron model, the laminar microcircuit
architecture, the hierarchical connectivity, the stimulation protocols, and
the data analysis pipeline.''

%==========================================================================
\subsection{Neuron Model and Simulation Environment}
%==========================================================================
All computational experiments were conducted using the NEST simulator, a
framework specifically optimized for the simulation of large-scale spiking
neural networks. The fundamental processing units of the network are modeled
as leaky integrate-and-fire (LIF) neurons with exact integration and
exponential-decaying postsynaptic currents. The subthreshold dynamics of the
membrane potential ($V_m$) are governed by the membrane time constant
($\tau_m = 10$ ms), membrane capacity ($C_m = 250$ pF), and a resting
potential ($V_{\text{rest}}$) of $-65$ mV. An action potential is generated
when $V_m$ reaches a critical threshold ($V_{\text{th}} = -50$ mV), after
which the potential is strictly clamped to a reset value
($V_{\text{reset}} = -65$ mV) for the duration of an absolute refractory
period ($\tau_{\text{ref}} = 2$ ms). Synaptic parameters were carefully tuned
to ensure network stability and to sustain the asynchronous-irregular firing
regimes characteristic of awake cortical states. Specifically, excitatory
synaptic weights were set to $w = 87.8 \pm 8.8$ pA with a decay time constant
$\tau_{\text{syn}} = 0.5$ ms, and the relative inhibitory synaptic strength
($g$) was fixed at $-4$. Transmission delays were uniformly distributed at
$1.5 \pm 0.75$ ms for excitatory synapses and $0.75 \pm 0.38$ ms for
inhibitory synapses. To emulate background cortical activity, each neuron
received external inputs modeled as Poisson spike trains with a mean rate of
8 Hz.

%==========================================================================
% TODO: ECUACIÓN DEL LIF
%==========================================================================
% [AQUÍ FALTA: Incluir la ecuación del potencial de membrana del LIF. Tienes
% dos opciones:
%   - Opción A (ecuación diferencial):
%     \tau_m \frac{dV_m}{dt} = -(V_m - V_{rest}) + R_m I(t)
%   - Opción B (formulación de conductancia con corrientes sinápticas):
%     \tau_m \frac{dV_m}{dt} = -(V_m - V_{rest}) + \sum_{j} w_j e^{-t/\tau_{syn}}
%
``Formally, the membrane potential evolves according to:
\begin{equation}
  \tau_m \frac{dV_m}{dt} = -(V_m - V_{\text{rest}}) + R_m I_{\text{syn}}(t),
\end{equation}
where $R_m$ is the membrane resistance and $I_{\text{syn}}(t)$ represents
the total synaptic current. Each synaptic input generates a
conductance-based current with exponential decay:
\begin{equation}
  I_{\text{syn}}(t) = w \, e^{-t / \tau_{\text{syn}}},
\end{equation}
where $w$ is the synaptic weight and $\tau_{\text{syn}}$ is the synaptic
time constant.''

%==========================================================================
\subsection{Laminar Microcircuit Architecture}
%==========================================================================
The primary visual cortex (V1) module is designed to represent a
$1 \text{ mm}^2$ patch of cortical tissue, based on the full-scale canonical
microcircuit model proposed by \citet{Potjans2014}. It comprises approximately
77,000 neurons distributed across four distinct laminar layers: L2/3, L4, L5,
and L6. Within each layer, the network maintains a physiological balance of
specific excitatory and inhibitory populations, approximating an 80/20 ratio.
Local recurrent connectivity is established using empirically derived
connection probabilities.

To capture intra-areal spatial integration, long-range horizontal projections
within V1 are incorporated. Crucially, these lateral connections are restricted
to the superficial L2/3 layer and are functionally segregated based on
distance. Nearby connections link functionally similar adjacent columns
(iso-orientation) with both excitatory and inhibitory synapses (transmission
delay $d_{\text{n.lat}} = 5 \pm 1.5$ ms). In contrast, distant connections
link columns with disparate orientation preferences and operate via a
predominantly suppressive mechanism (disynaptic inhibition) with a slower
conduction delay ($d_{\text{d.lat}} = 8 \pm 2.5$ ms).

%==========================================================================
% TODO: TABLA DE CONEXIONES O FIGURA DEL MICROCIRCUITO
%==========================================================================
% [AQUÍ FALTA: Referenciar una tabla o figura que resuma las probabilidades
% de conexión entre poblaciones. Si no quieres incluir la tabla completa de
% Potjans & Diesmann, al menos menciona que usas las mismas probabilidades y
% remite a una tabla en el apéndice.
%
``The complete connection probability matrix between the eight neuronal
populations (four layers $\times$ two types) follows the values reported by
\citet{Potjans2014} and is provided in Table~\ref{tab:connectivity} of the
Appendix for completeness.''

%==========================================================================
\subsection{Hierarchical Integration and Competitive Dynamics}
%==========================================================================
To investigate the mechanistic origins of extra-classical receptive field
(eCRF) effects, the baseline architecture was extended to include a
higher-order visual area (V2). The V2 module was adapted to mirror the
specific cytoarchitectural variations in laminar thickness and neuronal
densities of the macaque V2 region. Inter-areal communication is mediated by
targeted topographic projections operating with fast transmission delays
($2 \pm 0.5$ ms). Feedforward (FF) pathways originate from the superficial
layers (L2/3) of V1 and terminate predominantly in L4 and L6 of V2.
Conversely, top-down feedback (FB) projections originate from the deep layers
(L5 and L6) of V2 and target the superficial (L2/3) and infragranular (L5)
layers of V1, anatomically avoiding layer 4 to provide non-linear modulation
rather than driving inputs.

Crucially, to test the hypothesis that eCRF phenomena emerge from inter-areal
interactions rather than purely local filtering, a lateral competition
mechanism was embedded within the V2 module. This architecture utilizes two
spatially distinct, feature-aligned V2 modules that engage in lateral
competition mediated by strong, reciprocal inhibitory connections. By
enforcing this hierarchical winner-take-all competition, the top-down feedback
signals are non-linearly refined before modulating V1 activity. This
architectural feature is designed to overcome the limitations of simple linear
feedback models, ultimately reproducing the robust, stimulus-specific surround
suppression observed in physiological recordings.

%==========================================================================
% TODO: DESCRIPCIÓN FORMAL DE LAS TRES ARQUITECTURAS
%==========================================================================
% [AQUÍ FALTA: Describir explícitamente las tres configuraciones que
% comparaste, de forma que quede claro qué tiene cada una y qué no.
%
``To isolate the contribution of each circuit component, we systematically
compared three architectural configurations (see Figure~\ref{fig:architectures}):
\begin{description}
  \item[Configuration 1 (V1 lateral only):] Four V1 microcolumns arranged
        in a ring, interconnected exclusively via horizontal connections
        within L2/3. No V2 modules are present.\footnote{Se puede poner esto como pie de figura en lugar de aquí.}
  \item[Configuration 2 (V1 + linear V2 feedback):] The V1 columns project
        feedforward to a \emph{single} V2 module, which pools inputs from all
        columns and returns non-specific feedback to V1. This configuration
        implements hierarchical feedback without lateral competition.
  \item[Configuration 3 (V1 + competitive V2 feedback):] The V1 columns
        project feedforward to \emph{two} V2 modules, each aligned to a
        specific orientation column. Critically, the two V2 modules are
        connected via strong, reciprocal inhibitory synapses, implementing
        a winner-take-all competition.
\end{description}
All other parameters --- neuron counts, local connectivity, stimulation
protocols --- are held constant across configurations to ensure that
differences in suppression magnitude can be attributed exclusively to the
architectural manipulation.''

%==========================================================================
% TODO: FIGURA DE LAS TRES ARQUITECTURAS
%==========================================================================
% [AQUÍ FALTA: Referencia a una figura esquemática mostrando las tres
% configuraciones lado a lado. Esto es crítico para NeurIPS.
%
``A schematic of the three configurations is provided in
Figure~\ref{fig:architectures}.''

%==========================================================================
\subsection{Sensory Stimulation Protocols}
%==========================================================================
Bottom-up sensory stimulation was emulated through thalamocortical afferents
originating from the lateral geniculate nucleus (LGN), projecting primarily to
layers L4 and L6 of V1. Thalamic inputs were modeled as Poisson spike trains.
To simulate the classical receptive field (CRF) response, the thalamic firing
rate directed at the central microcolumn was increased to 20 Hz (simulating
optimal stimulus contrast) after a 500 ms baseline period. To evaluate the
extra-classical modulations, a dual asynchronous stimulation paradigm was
employed: following the central CRF activation at 500 ms, a distant peripheral
column was stimulated at 1000 ms. The intensity of this peripheral contextual
stimulus was systematically varied (10, 20, and 30 Hz) to quantify the
magnitude of surround suppression.

%==========================================================================
% TODO: DEFINICIÓN FORMAL DEL SUPPRESSION INDEX
%==========================================================================
% [AQUÍ FALTA: Definir matemáticamente el índice de supresión. Fundamental
% para que los resultados sean reproducibles.
%
``To quantify the strength of surround suppression, we define the
suppression index (SI) as:
\begin{equation}
  \text{SI} = 1 - \frac{R_{\text{center+surround}}}{R_{\text{center}}},
\end{equation}
where $R_{\text{center}}$ is the mean firing rate of the central column
during CRF stimulation alone, and $R_{\text{center+surround}}$ is the mean
firing rate during concurrent center and peripheral stimulation. An SI of 1
indicates complete suppression, while an SI of 0 indicates no contextual
modulation.''

%==========================================================================
\subsection{Data Analysis and Local Field Potentials}
%==========================================================================
Network dynamics and mean firing rates were evaluated using peri-stimulus
time histograms (PSTH). To connect the spiking network model to macroscopic
electrophysiological biomarkers --- such as gamma-band oscillations and
excitation/inhibition (E/I) balance alterations --- we simulated the Local
Field Potential (LFP) using a spatiotemporal kernel-based approximation. In
this framework, the macroscopic LFP was reconstructed as the linear
superposition of unitary field potentials (uLFP) generated by individual
action potentials, enabling detailed spectral analysis (PSD) of the network's
rhythmic activity under distinct contextual modulations.

%==========================================================================
% TODO: DETALLES DEL CÁLCULO DEL LFP Y ANÁLISIS ESPECTRAL
%==========================================================================
% [AQUÍ FALTA: Dar más detalles sobre el método de LFP. Los revisores de
% NeurIPS valoran la reproducibilidad. Debes cubrir:
%   - Referencia al método de Telenczuk o similar
%   - Ventana temporal y solapamiento para la PSD
%   - Método de estimación espectral (Welch, multitaper, etc.)
%
``Specifically, each spike at position $\mathbf{r}_i$ and time $t_k$
contributes a unitary field potential of the form
$\phi(\mathbf{r}, t) = K(|\mathbf{r} - \mathbf{r}_i|) \cdot h(t - t_k)$,
where $K(\cdot)$ is a spatial kernel modeling the extracellular decay and
$h(\cdot)$ is a temporal waveform representing the postsynaptic potential
\citep{Telenczuk2020}. Power spectral density (PSD) was estimated using
Welch's method with 500 ms Hamming windows and 50\% overlap. Gamma-band
power was quantified as the integrated PSD in the 30--100 Hz range.''

%==========================================================================
% TODO: COMPLEJIDAD COMPUTACIONAL Y RECURSOS
%==========================================================================
% [AQUÍ FALTA: Una frase o dos sobre los recursos computacionales usados.
% En NeurIPS esto es cada vez más esperado.
%
``All simulations were executed on a high-performance computing cluster
using [X] nodes with [Y] cores and [Z] GB of RAM. A typical simulation of
5 seconds of biological time for Configuration~3, comprising approximately
[N] neurons and [M] synapses, required roughly [T] hours of wall-clock time.
The code to reproduce all simulations and analyses will be released upon
publication.''